Formalizing Stack Safety as a Security Property

### 2023-01-05 submission5/core.tex
Tweaking the connection to lazy tagged enforcement, as we discuessed in the meeting Tuesday

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more generic characterization to the reader's intuition.  But these
mechanisms can be quite intricate, and the behaviors they aim to prevent are
subtle and varied. It can be hard to guess, based only on informal
intuitions, whether a given mechanism actually blocks all potential
attacks in a given class---or, conversely, whether it is overly conservative and disallows
system behaviors that are actually safe.
To settle such questions with confidence,
we need a precise, generic, and formal specification for stack
safety, as a basis both for comparing the security claims of different
enforcement techniques and for validating that these claims
are met by particular implementations. With the notable exception of the model used
in Skorstengaard et al.'s StkTokens work~\cite{SkorstengaardSTKJFP}, which we
discuss in Section \ref{sec:relwork}, there is a dearth of such properties in
the literature. \bcp{can we give a very quick preview here of the comparison?}

@@

@@

>>>>

The protections offered by such mechanisms are commonly described by giving
concrete examples of attacks that they can prevent---corruption of return
addresses, buffer overflows, use of uninitialized variables, etc.---leaving
a more formal characterization to the reader's intuition. Previous attempts
to produce a formal characterization do so using a {\it fully abstract overlay semantics},
a style of correct-by-construction machine \cite{SkorstengaardSTKJFP}.
The latest stack-safety overlay semantics \cite{Georges22:TempsDesCerises} capture
both control-flow and data protection (confidentiality and integrity).
But, this approach is not suitable for {\em lazy} enforcement mechanisms like those
of Roessler and DeHon, in which stores are not checked, and illegal stores are
only caught when their location is later read. We aim to provide a stack-safety
property suitable for testing lazy enforcement.

@@

Stack safety is challenging to specify in full generality, because the stack is fundamentally
an architecture-specific construct. Its desired behavior varies based on ISA, calling convention,
and the features of whichever programming language originated the code that must be protected.
Our approach centers around a tagged architecture, like PIPE \cite{???} or STAR \cite{???},
particularly using a lazy micro-policy. However, our design is factored into three pieces:
(1) the definition of {\em security relevant operations} in a given machine,
(2) the definition of a {\em security context} which tracks the security needs of different
elements of the state over the course of execution, and
(3) the definition, for a given security context, of how we distinguish secure behavior
from insecure. This separation of concerns allows (3) to be relatively general, while
architecture-specifics are isolated in (1) and some special features require extending
(2).

@@

\sna{Version that is less tied to lazy enforcement. It's still rather tied, actually,
  more than I remembered.}

One or more elements of a given state are {\it safe} if their values
in that state have no observable influence on future execution. We define this
formally as a sort of non-interference: for a state \(\sigma\) and
a set of elements \(\components\), \(\mathit{safe} ~ \sigma ~ \components\)
holds if we can replace the contents of \(\components\) with arbitrary values
and have the resulting state produce an equivalent trace. The state created
by modifying \(\components\) in this way is termed a variant.}
